Supplement to ``{\mytitle}''

### 2022-09-28 si/si.Rnw

registerDoRNG(18599687)
foreach(
  i = 1:length(lee1_epi_mif),
  .combine = rbind
) %dopar% {
  guess <- coef(lee1_epi_mif)
  sims <- simulate(h1_epi, nsim = N_LEE_SIMS, format = 'data.frame',
                   params = guess)  # 20 sims for each set of parameters
  sims$param_set <- i
  sims
} -> all_sims

quants <- all_sims %>%
  select(.id, week, cases) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

@@

<<PlotEpiDist, echo=FALSE, fig.cap="Compare to Figure S8 of Lee et al. (2020) supplement.", fig.width=7, fig.height=7, fig.align='center'>>=
epi_results <- as.data.frame(t(coef(lee1_epi_mif)))
epi_results$logLik <- logLik(lee1_epi_mif)

GGally::ggpairs(
  filter(epi_results, logLik > -2200),
  columns = c("rho", "tau", "beta1", "nu", "logLik")
)
@

@@

<<FitEnd, echo=FALSE, message=FALSE, include=FALSE>>=
foreach(
  i = 1:length(lee1_end_mif),
  .combine = rbind
) %dopar% {
  guess <- coef(lee1_end_mif)
  sims <- simulate(
    lee1_end_mif[[i]], nsim = N_LEE_SIMS,
    format = 'data.frame',
    params = guess
  )
  sims$param_set <- i
  sims
} -> end_sims

# quants_end <- end_sims %>%
#   select(.id, week, cases) %>%
#   group_by(week) %>%
#   summarize(
#     q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
#     q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
#     q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
#   ) %>%
#   ungroup() %>%
#   mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))
@

<<plotEndParams, fig.width=7, fig.height=7, fig.align='center', fig.cap="Bivariate relationships between variables after fitting endemic period. Compare to S9 of Lee et al. (2020) supplement.", cache=TRUE, echo=FALSE>>=

end_results <- as.data.frame(t(coef(lee1_end_mif)))
end_results$logLik <- logLik(lee1_end_mif)

GGally::ggpairs(
  filter(end_results, logLik > -2200),
  columns = c("rho", "tau", "beta1", "nu", "logLik")
)
@

@@

<<plotMod1Sims, echo=FALSE, fig.cap="Simulations from various parameter sets. Simulations from parameter sets with likelihood $<$ -2200 are removed. Compare to Figure S7 of \\cite{lee20sup}.", cache=TRUE, fig.width=6.5, fig.height=3.75>>=
end_complete <- filter(end_results, logLik > -2200)
bad_end_pars <- which(end_results$logLik < -2200)

quants_end <- end_sims %>%
  filter(!param_set %in% bad_end_pars) %>%
  select(.id, week, cases) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

@@

% \renewcommand\missing{---}
%
% \begin{table}
% \begin{center}
% \begin{tabular}{|l|c|c|c|c|}
% \hline
% Mechanism & Unified & Old model 1 & Old model 2 & Old model 3
% \\
% \hline
% \hline
% Recovery rate
%   & $\muIR$
%   & $\gamma$
%   & $\gamma$
%   & $\gamma$
% \\
% Latency
%   & $\muEI$
%   & $\sigma$
%   & $\sigma$
%   & \missing
% \\
% Direct transmission
%   & $\transmission$
%   & $\beta_{1:6}$
%   & $\beta$
%   & $c^i$
% \\
% Rainfall exponent
%   & \rainfallExponent
%   & \missing
%   & \missing
%   & $r$
% \\
% Lost immunity
%   & $\muRS$
%   & $\alpha$
%   & $\alpha$
%   & $\rho$
% \\
%   & $\omega_1$, $\omega_2$
%   & \missing
%   & $\omega_1$, $\omega_2$
%   & \missing
% \\
% Vaccine efficacy
%   & $\vaccineEfficacy_{1:2}$
%   & \missing
%   & $\theta_{1:2}$
%   & \missing
% \\
% Birth rate
%   & $\muBirth$
%   & $\mu$
%   &
%   & $\mu$
% \\
% Death rate
%   & $\muDeath$
%   & $\delta$
%   &
%   & $\mu$
% \\
% Symptomatic frac.
%   & $\symptomFrac$
%   & $\theta_k(t)=c\theta^*(t-\tau_k)$
%   & $k$
%   & $\theta$
% \\
% Asympt.~infectivity
%   & $\asymptomRelativeInfect$
%   & $(1-\kappa)$
%   & $\beta_A/\beta$
%   & 1
% \\
% Asympt.~shedding
%   & $\asymptomRelativeShed$
%   & \missing
%   & $\mu_A/\mu$
%   & $\theta_A/\theta_I$
% \\
% Seasonality amplitude
%   & $\seasAmplitude$
%   & \missing
%   & $\alpha_s$
%   & $a$
% \\
% Water to human
%   & $\Wbeta{}$
%   & \missing
%   & $\beta_W$
%   & $\beta^i$
% \\
%   & $\Wsat$
%   & \missing
%   & $\mathrm{sat}$
%   & $1$
% \\
% Cholera decay in water
%   & $\Wremoval$
%   & \missing
%   & $\delta$
%   & $\mu_B$
% \\
% Human to water
%   & $\Wshed$
%   & \missing
%   & $\mu$
%   & $\theta_I$
% \\
% Mixing exponent
%   & $\mixExponent$
%   & $\nu$
%   & \missing
%   & \missing
% \\
% Process noise
%   & $\sigmaProc$
%   & \missing
%   & \missing
%   & $\sigma$
% \\
% Reporting rate
%   & $\reportRate$
%   & $\rho$
%   & $\rho$
%   & $\epsilon$
% \\
% Observation overdispersion
%   & $\obsOverdispersion$
%   & $\tau$
%   & \missing
%   & $k$
% \\
% \hline
% \end{tabular}
% \end{center}
% \caption{Translation from the unified notation of this article back into the notations used for Lee et al.~(2020)}
% \label{tab:translate}
% \end{table}

% \section{Investigation into model 2}
%
% As part of our case study into the Haiti paper, model 2 was recreated using the spatPomp package in R. To confirm the validity of our model, we recreated Figure S15 from the supplement. In the original model, the authors reinitialized the state values for the endemic phase starting on March 2014. In particular, the group assumed that 75\% of the population was susceptible to cholera. We believe this choice to be unnatural. Having the state values change in the middle of a simulation has no causal interpretation. Instead, the most logical thing to do when fitting an epidemic and endemic model would be to carry over the state values from the former to the latter. This keeps the integrity of the model intact.
%
% In Figure S-2, we created the same endemic model with the only difference being the state values are carried over from the end of the epidemic phase. Here, we see that the model predicts a much different trajectory than in Figure S-1. This suggests that the decision to reinitialize state values in the middle of a simulation, as in the original model, is likely a poor choice since it has no cause and does not reflect the nature of the epidemic. One consideration is the choice of measurement component. In the original model, we use a linear measurement component. To avoid a situation where the trajectory goes to zero, we can implement a logarithmic measurement component. This is a more natural measure since the fit should be relative to the magnitude of the reported data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage

>>>>

registerDoRNG(18599687)
foreach(
  i = 1:length(lee1_epi_mif),
  .combine = rbind
) %dopar% {
  guess <- lee1_epi_mif$lee1_epi_mif_params[i, ] 
  sims <- simulate(h1_epi, nsim = N_LEE_SIMS, format = 'data.frame',
                   params = guess)  # 20 sims for each set of parameters
  sims$param_set <- i
  sims %>%
    dplyr::select(.id, week, cases, param_set)
} -> all_sims

@@

<<PlotEpiDist, echo=FALSE, fig.cap="Compare to Figure S8 of Lee et al. (2020) supplement.", fig.width=7, fig.height=7, fig.align='center'>>=
lee1_epi_evals <- readRDS(file = paste0("../model1/", RL_dir, "lee1_epi_evals.rds"))

epi_results <- as.data.frame(lee1_epi_mif$lee1_epi_mif_params)
epi_results$logLik <- lee1_epi_evals$pfLL

GGally::ggpairs(
  dplyr::filter(epi_results, logLik > -20000),
  columns = c("rho", "tau", "beta1", "nu", "logLik")
)
@

@@

<<FitEnd, echo=FALSE, message=FALSE, include=FALSE>>=
foreach(
  i = 1:nrow(lee1_end_mif),
  .combine = rbind
) %dopar% {
  r_params <- lee1_end_mif[i, ]

  lee1_epi_mif$start_states %>%
    filter(param_set == i) %>%
    select(-param_set) %>%
    sapply(., median, na.rm = TRUE) -> rinit_parms

  # Set the initial state
  new_rinit <- Csnippet(
    sprintf(
      "
          S = nearbyint(%f);
          E = nearbyint(%f);
          I = nearbyint(%f);
          A = nearbyint(%f);
          R = nearbyint(%f);
          incid = nearbyint(%f);
          foival = %f;
          Str0 = nearbyint(%f);
          Sout = nearbyint(%f);
          Sin = nearbyint(%f);
          ",
      rinit_parms['S'],
      rinit_parms['E'],
      rinit_parms['I'],
      rinit_parms['A'],
      rinit_parms['R'],
      rinit_parms['incid'],
      rinit_parms['foival'],
      rinit_parms['Str0'],
      rinit_parms['Sout'],
      rinit_parms['Sin']
    )
  )

  h1_end <- h1_end %>% pomp(
    rinit = new_rinit,
    statenames = c(
      "S", "E", "I", "A", "R", "incid",
      "foival", "Str0", "Sout", "Sin"
    )
  )

  coef(h1_end) <- r_params

  sims <- simulate(
    h1_end, nsim = N_LEE_SIMS,
    format = 'data.frame',
    params = r_params
  )
  sims$param_set <- i
  sims %>%
    dplyr::select(.id, week, cases, param_set)
} -> end_sims

<<plotEndParams, fig.width=7, fig.height=7, fig.align='center', fig.cap="Bivariate relationships between variables after fitting endemic period. Compare to S9 of Lee et al. (2020) supplement.", cache=TRUE, echo=FALSE>>=
lee1_end_evals <- readRDS(file = paste0("../model1/", RL_dir, "lee1_end_evals.rds"))

end_results <- as.data.frame(lee1_end_mif)
end_results$logLik <- lee1_epi_evals$pfLL

GGally::ggpairs(
  filter(end_results, logLik > -20000),
  columns = c("rho", "tau", "beta1", "nu", "logLik")
)
@

@@

<<plotMod1Sims, echo=FALSE, fig.cap="Simulations from various parameter sets. Simulations from parameter sets with likelihood $<$ -2200 are removed. Compare to Figure S7 of \\cite{lee20sup}.", cache=TRUE, fig.width=6.5, fig.height=3.75>>=

epi_complete <- filter(epi_results, logLik > -20000)
bad_epi_pars <- which(epi_results$logLik < -20000)

quants <- all_sims %>%
  filter(!param_set %in% bad_epi_pars) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

end_complete <- filter(end_results, logLik > -20000)
bad_end_pars <- which(end_results$logLik < -20000)

quants_end <- end_sims %>%
  filter(!param_set %in% bad_end_pars) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

@@

\clearpage
